Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2014-01-07 paper.Rnw
Adding WA xval perturb tests

The hippocampus is a brain structure situated in the medial temporal
lobe, and has long been associated with learning and memory
\citep{DenHeijer2012,Scoville2000}. The hippocampus is of interest to
clinical neuroscientists because it is implicated in many forms of brain
dysfunction, including Alzheimer's disease \citep{Sabuncu2011} and
schizophrenia \citep{Narr2004,Karnik-Henry2012}. In neuroimaging studies,
structural magnetic resonance images (MRI) are often used for the volumetric
assessment of the hippocampus.  As such, accurate segmentation of the
hippocampus and its subfields in MRI is a necessary first step to better
understand the inter-individual variability of subject neuroanatomy.

@@

However, many of these methods require significant investment of time and
resources for the creation of the atlas library ranging between 30
\citep{Heckemann2006} and 80 \citep{Collins2010} manually segmented atlases.
This strategy has the main drawback of being inflexible as it does not easily
accommodate varying the definition of the hippocampal anatomy (such as the
commonly used heuristic of subdividing the hippocampus into head, body, and tail
\citep{Poppenk2011,Pruessner2000}). Furthermore, none of these methods have
demonstrated sufficient flexibility to accommodate atlases that are somehow
exceptional such as those derived from serial histological data
\citep{Chakravarty2006,Yelnik2007} or high-resolution MRI data that enables
robust identification of hippocampal subfields
\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}. Due
to the recent availability of the latter, there has been increased interest in
the use of probabilistic methods for the identification of the hippocampal
subfields on standard T1-weighted images. Our group recently demonstrated that
through use of an intermediary automated segmentation stage, robust and accurate
segmentation of the striatum, pallidum, and thalamus using a single atlas
derived from serial histological data is possible \citep{Chakravarty2013}.
The novelty of this manuscript is the extension of our \multiatlas{} methodology
to the hippocampus using more than a single input atlas, while simultaneously
limiting the number of inputs used during segmentation, and demonstrating that
accurate identification of the hippocampal subfields is indeed possible using
this methodology.

@@

Also of interest to this manuscript are the methods that attempt to define
hippocampal subfields using standard T1- or T2-weighted data, of which there are
few.  \citet{VanLeemput2009} demonstrate that the applicability of hippocampal
subfield segmentation in T1-weighted images by using Markov random field priors
learned from 10 manual segmentations.  This work, available as part of the
\freesurfer{} package, is limited as the segmentation omits the tail of the
hippocampus and the protocol has yet to be fully validated.
\citet{Yushkevich2009} manually segment hippocampal subfields on high-resolution
(either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$ resolution
voxels) T2-weighted MR images acquired from five post-mortem medial temporal
lobe samples. Then, using nonlinear registration guided by shape-based models of
the subfield segmentations, and manually derived hippocampus masks of the target
images, the authors demonstrate accurate parcellation of hippocampal subfields
in clinical 3T T1-weighted MRI volumes.  Using \multiatlas with bias correction
techniques, \citet{Yushkevich2010} demonstrate a semi-automated method of
subfield segmentation on in vivo focal T2-weighted MR acquisitions of the
temporal lobe. Manual input is only needed to mark mark divisions between the
head, body and tail of the hippocampus on target images.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%        Experiment 2: Hippocampal Subfield Cross-Validation          %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
The previous experiment assesses \mb{} peformance on whole hippocampus
segmentation. In this experiment, we aim to evaluate \mb{} hippocampal subfield
segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
use a modified leave-one-out cross-validation (LOOCV) design.

@@

In this experiment, the Winterburn atlases are resampled to $0.9mm^3$-isotropic
voxel images to simulate standard 3T T1-weighted resolution images. Image
subsampling is performed using trilinear subsampling techniques.

In each round of LOOCV, one atlas is selected and segmented with \mb{} once
using each possible triplet of atlas images from the remaining four unselected
atlases. We chose to use an atlas library with an odd number of images so as to
ensure an unbiased label fusion when using majority voting (see Discussion).

@@

\begin{figure}
<<fig:WAxval-kappa>>=
wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
wa.n.sim$label = as.factor(wa.n.sim$label)
levels(wa.n.sim$label) <-list(
  "CA1"=1, "CA1"=101, "Subiculum"=2,"Subiculum"=102, 
  "CA4/DG"=4,"CA4/DG"=104, "CA2/CA3"=5,"CA2/CA3"=105, 
  "SR/SL/SM"=6,"SR/SL/SM"=106)

@@

Figure \ref{fig:WAval-kappa} shows the overlap similarity scores between the
\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
subregion across all subjects and folds of the validation. Mean (standard
deviation) DSC scores of the subregions are: 
  \Sexpr{mean(CA1)}   ($\pm$\Sexpr{sd(CA1)}) for the CA1 region,  
  \Sexpr{mean(CA23)}  ($\pm$\Sexpr{sd(CA23)}) for the CA2/CA3 regions,  
  \Sexpr{mean(CA4DG)} ($\pm$\Sexpr{sd(CA4DG)}) for the CA4/DG regions,  
  \Sexpr{mean(Sub)}   ($\pm$\Sexpr{sd(Sub)}) for the Subiculum region, and
  \Sexpr{mean(SRLM)}  ($\pm$\Sexpr{sd(SRLM)}) for the moleculare regions.
Bland-Altman plots comparing \mb{} to resampled subregion volumes reveal a
proportional bias towards underestimating volumes in all subregions (Figure
\ref{fig:WAxval-Bland-Altman}). Prominent fixed biases in the volumes of the
CA4/DG and SR/SL/SM subregions are congruent with the lower similarity scores
seen in these regions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
%
\begin{sidewaystable}[!tbp]
\scriptsize
\caption{\textbf{Automated segmentation accuracy of the Hippocampus} 
This table shows the best published Dice's overlap measure between automated and
ground truth segmentations of various methods.  Unless otherwise specified,
validation datasets are composed equally of cases and control subjects, and use
manual segmenation labels as ground truth in computing DSC scores. AD =
Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
(CN); LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
cross-validation; SNT = Surgical Medtronic Navigation Technologies
semi-automated labels.  Some studies of automated segmentation of ADNI images
are excluded because they do not not provide overlap measures for the
hippocampus \citep{Heckemann2011,Chupin2009}.
  \label{tab:other-methods}} 
\begin{center}
\begin{tabular}{ p{1.2in}  c  p{1.1in} l  p{3.2in} p{1in} }
\hline\hline
Method & Atlases & DSC \newline {\tiny mean (AD; MCI; CN)} & Reference & Validation & Dataset (Truth)\tabularnewline
\hline
%%  
\mb{}                     &9 & \Sexpr{snt.mb.kappa.best} & \cite{Pipitone2014}
& 10-fold MCCV on 69 subjects 
& ADNI (SNT) 
\tabularnewline 
%%  
Patch-based label fusion  &16&0.861 {\tiny (0.838; ---; 0.883)} &\cite{Coupe2011e}
& LOOCV 
& ADNI (SNT) 
\tabularnewline
%%  
Multi-atlas               &20&0.848 {\tiny (---; 0.798, 0.898)} &\cite{Wang2011}  
& 10-fold MCCV on 20 of 139 subjects
& ADNI (SNT) 
\tabularnewline
%%  
ACM (AdaBoost-based)      &21&0.862                           &\cite{Morra2008} 
& LOOCV                                          
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
LEAP                      &30&0.848                           &\cite{Wolz2010}  
& Segmentation of 182 subjects                   
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
Multi-atlas               &30&0.885                           &\cite{Lotjonen2010}
& Segmentation of 60 subjects                    
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
Multi-atlas (MAPS)        &55&0.890                           &\cite{Leung2010}
& Segmentation of 30 subjects (10 AD, MCI, and CN)
& ADNI (SNT) 
\tabularnewline

@@

%%
\mb{}                      &9 & \Sexpr{mb.kappa.best}       &\cite{Pipitone2014} 
& 10-fold MCCV on 60 subjects 
& ADNI (\prsnr)
\tabularnewline 
%%
Neural nets                &10&0.740   &\cite{Powell2008} % {\tiny (0.72, 0.75)} 
& Segmentation of 5 subjects 
& controls %Other
\tabularnewline
%%
Probablistic atlas         &11&0.852   &\cite{VanderLijn2008}
& 11 atlases used in 100 rounds of LOOCV on 20 elderly subjects
& elderly controls %Other
\tabularnewline
%%
Probabilistic Atlas        &16&0.860   &\cite{Chupin2009}
& LOOCV 
& AD subjects %Other
\tabularnewline
%%
Anatomically-guided EM     &17&0.812   &\cite{Pohl2007} %{\tiny (0.815, 0.808)} 
& LOOCV on 17 controls, segmentation of 33 mixed subjects
& mixed diagnosis %Other
\tabularnewline
%%
Multi-atlas                &30&0.820   &\cite{Heckemann2006} %{\tiny (0.83, 0.81)}    
& LOOCV
& controls %Other
\tabularnewline
%%
Multi-atlas                &30&0.880   &\cite{Gousias2008} %{\tiny (0.88,---)}    
& 30 adult atlas used, segmentation of 33 2yr old subjects 
& 2yr old controls  %Other
\tabularnewline
%%
Multi-atlas                &80&0.890   &\cite{Collins2010}
& LOOCV                 
& controls %Other       
\tabularnewline
%%                      
Multi-atlas                &55&0.860   &\cite{Barnes2008}%{\tiny (---,0.86)}    
& LOOCV 
& controls and AD %Other
\tabularnewline
%%
Multi-atlas               &275&0.835   &\cite{Aljabar2009}% {\tiny (0.832, 0.837)} 
& LOOCV 
& controls %Other
\tabularnewline
\hline
\end{tabular}
\end{center}
\end{sidewaystable}

@@

Finally, we have provided evidence that using the Winterburn high-resolution
hippocampal subfield atlases \citep{Winterburn2013} our algorithmic framework is
appropriate for the segmentation of hippocampal subfields in standard
T1-weighted data. Subfield segmentation is a burgeoning topic in the literature
although very few automated methods are available for the segmentation of 3T
data \citep{Yushkevich2009,VanLeemput2009}. While recent work demonstrates that
subfield segmentations can be used for classification of AD, MCI, and CN, there
has been no explicit validation of the segmentations being produced other than
indirectly via classification accuracy (for instance, the semi-automated
subfield segmentations produced in \citet{Yushkevich2009}).  Our work
demonstrates that we can reliably reproduce segmentations for the CA1,
subiculum, and CA4/dentate subregions.  That the CA2/CA3 and molecular
layers are less well reproduced should not be surprising as these are extremely
thin and spatially convoluted regions that originally required high-resolution
MRI for identification and so it is likely that the extents of these regions are
well below the resolution and contrast offered by standard T1-weighted images.

>>>>

The hippocampus is a brain structure situated in the medial temporal
lobe, and has long been associated with learning and memory
\citep{DenHeijer2012,Jeneson2012,Wixted2011,Scoville2000}. The hippocampus is of
interest to clinical neuroscientists because it is implicated in many forms of
brain dysfunction, including Alzheimer's disease \citep{Sabuncu2011} and
schizophrenia \citep{Narr2004,Karnik-Henry2012}. In neuroimaging studies,
structural magnetic resonance images (MRI) are often used for the volumetric
assessment of the hippocampus.  As such, accurate segmentation of the
hippocampus and its subfields in MRI is a necessary first step to better
understand the inter-individual variability of subject neuroanatomy.

@@

However, many of these methods require significant investment of time and
resources for the creation of the atlas library ranging between 30
\citep{Heckemann2006} and 80 \citep{Collins2010} manually segmented atlases.
This strategy has the main drawback of being inflexible as it does not easily
accommodate varying the definition of the hippocampal anatomy (such as the
commonly used heuristic of subdividing the hippocampus into head, body, and tail
\citep{Poppenk2011,Pruessner2000}). Furthermore, none of these methods have
demonstrated sufficient flexibility to accommodate atlases that are somehow
exceptional such as those derived from serial histological data
\citep{Chakravarty2006,Yelnik2007} or high-resolution MRI data that enables
robust identification of hippocampal subfields
\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}. Due
to the recent availability of the latter, there has been increased interest in
the use of probabilistic methods for the identification of the hippocampal
subfields on standard T1-weighted images. Our group recently demonstrated that
through use of an intermediary automated segmentation stage, robust and accurate
segmentation of the striatum, pallidum, and thalamus using a single atlas
derived from serial histological data is possible \citep{Chakravarty2013}.
The novelty of this manuscript is the extension of our \multiatlas{} methodology
to the hippocampus using more than a single input atlas, while simultaneously
limiting the number of inputs used during segmentation, and so demonstrating that
reliable application of the atlases' subfield segmentation to an unlabelled
target image is indeed possible using this methodology.

@@

Also of interest to this manuscript are the methods that attempt to define
hippocampal subfields using standard T1- or T2-weighted data, of which there are
few.  \citet{VanLeemput2009} demonstrate that the applicability of hippocampal
subfield segmentation in T1-weighted images by Bayesian techniques using Markov
random field shape priors learned from 10 manual segmentations.  This work,
available as part of the \freesurfer{} package, is limited as the segmentation
omits the tail of the hippocampus and the protocol has yet to be fully
validated.  \citet{Yushkevich2009} manually segment hippocampal subfields on
high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
resolution voxels) T2-weighted MR images acquired from five post-mortem medial
temporal lobe samples. Then, using nonlinear registration guided by shape-based
models of the subfield segmentations, and manually derived hippocampus masks of
the target images, the authors demonstrate accurate parcellation of hippocampal
subfields in clinical 3T T1-weighted MRI volumes.  Using \multiatlas with bias
correction techniques, \citet{Yushkevich2010} demonstrate a semi-automated
method of subfield segmentation on in vivo focal T2-weighted MR acquisitions of
the temporal lobe. Manual input is only needed to mark mark divisions between
the head, body and tail of the hippocampus on target images.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%        Experiment 2: Hippocampal Subfield Cross-Validation          %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
The previous experiment assesses \mb{} peformance on whole hippocampus
segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
use a modified leave-one-out cross-validation (LOOCV) design.

@@

In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
voxel images to simulate standard 3T T1-weighted resolution images. Image
subsampling is performed using trilinear subsampling techniques.  In each round
of LOOCV, a single atlas is selected and segmented with \mb{} once using each
possible triplet of atlas images from the remaining four unselected atlases.
Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
with an odd number of images so as to ensure an unbiased label fusion when using
majority voting (see Discussion).

@@

\begin{figure}
<<fig:WAxval-kappa>>=
wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
wa.n.sim$label = as.factor(wa.n.sim$label)
levels(wa.n.sim$label) <-list(
  "CA1"=1, "CA1"=101, 
  "CA2/CA3"=5,"CA2/CA3"=105, 
  "CA4/DG"=4,"CA4/DG"=104, 
  "Subiculum"=2,"Subiculum"=102, 
  "SR/SL/SM"=6,"SR/SL/SM"=106)

@@

Figure \ref{fig:WAval-kappa} shows the overlap similarity scores between the
\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
subregion across all subjects and folds of the validation. Mean and standard
deviation DSC scores of the subregions are: 
  \Sexpr{mean(CA1)}  $\pm$\Sexpr{sd(CA1)} for the CA1 region,  
  \Sexpr{mean(CA23)} $\pm$\Sexpr{sd(CA23)} for the CA2/CA3 regions,  
  \Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)} for the CA4/DG regions,  
  \Sexpr{mean(Sub)}  $\pm$\Sexpr{sd(Sub)} for the Subiculum region, and
  \Sexpr{mean(SRLM)} $\pm$\Sexpr{sd(SRLM)} for the moleculare regions.

Bland-Altman plots comparing \mb{} to resampled subregion volumes reveal a
proportional bias towards underestimating volumes in all subregions (Figure
\ref{fig:WAxval-Bland-Altman}). Prominent fixed biases in the volumes of the
CA4/DG and SR/SL/SM subregions are congruent with the lower similarity scores
seen in these regions.

<<tab:WAxval-perturb-kappa, echo=F,results="asis">>=
wa.perturb = read.csv('data/winterburn.perturb.kappa.csv')
wa.perturb = wa.perturb[c("subject","label","operation","k")]
wa.perturb$label = as.factor(wa.perturb$label)
levels(wa.perturb$label) <-list(
  "CA1"="1", "CA1"="101",  "CA2/CA3"="5","CA2/CA3"="105",
  "CA4/DG"="4","CA4/DG"="104",  "Subiculum"="2","Subiculum"="102", 
  "SR/SL/SM"="6", "SR/SL/SM"="106")
mean_range = function (v) {
  sprintf("$%0.2f \\pm %0.2f$", mean(v), sd(v))
}
wa.perturb.table = dcast(wa.perturb, label ~ operation, fun.aggregate=mean_range)
wa.sim.sum = dcast(wa.n.sim, label ~ ., value.var = 'k', fun.aggregate=mean_range)
names(wa.sim.sum)[2] = 'MAGeT'
wa.perturb.table = merge(wa.perturb.table, wa.sim.sum)

wa.perturb.table=wa.perturb.table[c('label','MAGeT', 'translate','contract','expand')]
names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")

caption="Overlap similarity results for the each of the subfields of the hippocampus. 
Simulated overlap similarity results are also given for resampled labels that were 
translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
Values are given as mean Dice's Kappa coefficient \\pm stddev."
latex(wa.perturb.table, #file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      label="tab:WAxval-perturb-kappa", rowname=NULL)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
%
\begin{sidewaystable}[!tbp]
\scriptsize
\caption{\textbf{Automated segmentation accuracy of the Hippocampus} 
This table shows the best published Dice's overlap measure between automated and
ground truth segmentations of various methods.  Unless otherwise specified,
validation datasets are composed equally of cases and control subjects, and use
manual segmenation labels as ground truth in computing DSC scores. AD =
Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
(CN); LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
cross-validation; SNT = Surgical Medtronic Navigation Technologies
semi-automated labels.  Some studies of automated segmentation of ADNI images
are excluded because they do not not provide overlap measures for the
hippocampus \citep{Heckemann2011,Chupin2009}.
  \label{tab:other-methods}} 
\begin{center}
\begin{tabular}{ p{1.2in}  c  p{1.1in} l  p{3.2in} p{1in} }
\hline\hline
Method & Atlases & DSC \newline {\tiny mean (AD; MCI; CN)} & Reference & Validation & Dataset (Truth)\tabularnewline
\hline
%%  
\mb{}                     &9 & \Sexpr{snt.mb.kappa.best} &
& 10-fold MCCV on 69 subjects 
& ADNI (SNT) 
\tabularnewline 
%%  
Patch-based label fusion  &16&0.861 {\tiny (0.838; ---; 0.883)} &\cite{Coupe2011e}
& LOOCV 
& ADNI (SNT) 
\tabularnewline
%%  
Multi-atlas               &20&0.848 {\tiny (---; 0.798, 0.898)} &\cite{Wang2011}  
& 10-fold MCCV on 20 of 139 subjects
& ADNI (SNT) 
\tabularnewline
%%  
ACM (AdaBoost-based)      &21&0.862                           &\cite{Morra2008} 
& LOOCV                                          
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
LEAP                      &30&0.848                           &\cite{Wolz2010}  
& Segmentation of 182 subjects                   
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
Multi-atlas               &30&0.885                           &\cite{Lotjonen2010}
& Segmentation of 60 subjects                    
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
Multi-atlas (MAPS)        &55&0.890                           &\cite{Leung2010}
& Segmentation of 30 subjects (10 AD, MCI, and CN)
& ADNI (SNT) 
\tabularnewline

@@

%%
\mb{}                      &9 & \Sexpr{mb.kappa.best}       &
& 10-fold MCCV on 60 subjects 
& ADNI (\prsnr)
\tabularnewline 
%%
Neural nets                &10&0.740   &\cite{Powell2008} % {\tiny (0.72, 0.75)} 
& Segmentation of 5 subjects 
& controls %Other
\tabularnewline
%%
Probablistic atlas         &11&0.852   &\cite{VanderLijn2008}
& 11 atlases used in 100 rounds of LOOCV on 20 elderly subjects
& elderly controls %Other
\tabularnewline
%%
Probabilistic Atlas        &16&0.860   &\cite{Chupin2009}
& LOOCV 
& AD subjects %Other
\tabularnewline
%%
Anatomically-guided EM     &17&0.812   &\cite{Pohl2007} %{\tiny (0.815, 0.808)} 
& LOOCV on 17 controls, segmentation of 33 mixed subjects
& mixed diagnosis %Other
\tabularnewline
%%
Multi-atlas                &30&0.820   &\cite{Heckemann2006} %{\tiny (0.83, 0.81)}    
& LOOCV
& controls %Other
\tabularnewline
%%
Multi-atlas                &30&0.880   &\cite{Gousias2008} %{\tiny (0.88,---)}    
& 30 adult atlas used, segmentation of 33 2yr old subjects 
& 2yr old controls  %Other
\tabularnewline
%%
Multi-atlas                &80&0.890   &\cite{Collins2010}
& LOOCV                 
& controls %Other       
\tabularnewline
%%                      
Multi-atlas                &55&0.860   &\cite{Barnes2008}%{\tiny (---,0.86)}    
& LOOCV 
& controls and AD %Other
\tabularnewline
%%
Multi-atlas               &275&0.835   &\cite{Aljabar2009}% {\tiny (0.832, 0.837)} 
& LOOCV 
& controls %Other
\tabularnewline
\hline
\end{tabular}
\end{center}
\end{sidewaystable}

@@

Finally, we have provided evidence that using the Winterburn high-resolution
hippocampal subfield atlases \citep{Winterburn2013} our algorithmic framework is
appropriate for the segmentation of hippocampal subfields in standard
T1-weighted data. Subfield segmentation is a burgeoning topic in the literature
although very few automated methods are available for the segmentation of 3T
data \citep{Yushkevich2009,Yushkevich2010,VanLeemput2009}.  Table
\ref{tab:subfield-segmentation-kappa} compares segmentation accuracy from some
of these methods and \mb{}. The overlap DSC scores for \mb{} subregions are
notably lower but a direct comparison of overlap values must be done cautiously.
In the present work, our overlap scores are computed on $0.9mm$-isotropic voxel
resolution images, whereas \citet{Yushkevich2010} uses focal
$0.4\times0.5\times2.0mm$ voxel resolution images, and \citet{VanLeemput2009}
use supersampled $0.380mm$-isotropic voxel resolution images. The larger voxel
images we use necessarily entail a greater change in DSC for each incorrectly
labelled voxel.  In addition, our automated segmentations are compared to manual
segmentations resampled from $0.3mm$-isotropic voxel labels; the resampling
process inevitably introduces noise which may lower overlap scores.  Lastly, as
our method is aimed specifically at situations when manually produced atlases
are scarce, in our cross validation we are forced to use three rather than all
five of the Winterburn atlases which, based on our findings with whole
hippocampal segmentation, would lead to improved overlap similarity.
Nevertheless, the advantage of evaluating \mb{} on standard 3T T1-weighted
resolution MR images with a publically available atlas library is that our
results reflect typical usage scenarios of researchers and clinicians.

Our overlap DSC values demonstrates that we can reliably reproduce segmentations
for the CA1, subiculum, and CA4/dentate subregions.  That the CA2/CA3 and
molecular layers are less well reproduced (DSC < 0.5) should not be surprising
as these are extremely thin and spatially convoluted regions that originally
required high-resolution MRI for identification and so it is likely that the
extents of these regions are well below the resolution and contrast offered by
standard T1-weighted images.

\begin{table}[!tbp]
\scriptsize
\caption{\textbf{A comparison of subfield segmentation overlap similarity with
manual raters.} \label{tab:subfield-segmentation-kappa}} 
\begin{center}
\begin{tabular}{ l c c c }
\hline\hline
Subfield             & \mb{}               & \citet{VanLeemput2009} & \citet{Yushkevich2010}  \tabularnewline
\hline                                   
CA1                  &\Sexpr{mean(CA1)}    & 0.62 & 0.875                                     \tabularnewline
CA2/3                &\Sexpr{mean(CA23)}   & 0.74 & $CA2=0.538, CA3=0.618$                    \tabularnewline
CA4/DG               &\Sexpr{mean(CA4DG)}  & 0.68 & $DG=0.873$                                \tabularnewline
presubiculum         & ---                 & 0.68 & ---                                       \tabularnewline
subiculum            &\Sexpr{mean(Sub)}    & 0.74 & 0.770                                     \tabularnewline
hippocampal fissure  & ---                 & 0.53 & ---                                       \tabularnewline
SR/SL/SM             &\Sexpr{mean(SRLM)}   & ---  & ---                                       \tabularnewline
fimbria              & ---                 & 0.51 & ---                                       \tabularnewline
head                 & ---                 & ---  & 0.902                                     \tabularnewline
tail                 & ---                 & ---  & 0.863                                     \tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}
